---
title: "Solution set 3"
author: "Ankit Mathur"
date: "2/17/2019"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy = TRUE)
library(tidyverse)
library(GGally)
library(broom)
cb_palette = c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
```

# Solution 1

```{r, include=FALSE}
# Load the data
movies = read.csv("movie_budgets.txt", sep = " ")
```

```{r, include=FALSE}
# Transform movie budgets to log scale
movies$log_budget <- log10(movies$budget)
```

```{r, include=FALSE}
# Plot *ggpairs* for all 3 vars
ggpairs(movies, columns = c("log_budget", "length", "year"))
```

```{r, include=FALSE}
## Color trivariate plots by year
# unique(movies[order(movies$year), "year"])
# names(movies)
fac_year = rep(NA, nrow(movies))
fac_year[movies$year < 1960] = "1906-1960"
fac_year[movies$year >= 1960] = "1960-2005"
movies2 = data.frame(movies, fac_year)
ggpairs(movies2, columns = c("log_budget", "length", "year"), aes(color = fac_year))
```

```{r, include=FALSE}
# Color trivariate plots by length
fac_length = rep(NA, nrow(movies))
fac_length[movies$length < 90] = "short"
fac_length[(movies$length >= 90) & (movies$length < 120)] = "medium"
fac_length[movies$length >= 120] = "long"
movies3 = data.frame(movies, fac_length)
ggpairs(movies3, columns = c("log_budget", "length", "year"), aes(color = fac_length))
```


```{r, include=FALSE}
# Plot budget against length faceted by year
ggplot(movies2, aes(x = length, y = log_budget)) + geom_point() +
  geom_smooth(method = "loess", span = 0.8, method.args = list(degree = 1)) +
  # facet_wrap(~ cut_number(year, n = 3), ncol = 2) +
  facet_wrap(~ fac_year) +
  labs(title = "Movie budgets vs. lengths cut by year")
```

```{r, include = FALSE}
# Plot budget against year faceted by length
ggplot(movies3, aes(x = year, y = log_budget)) + geom_point() +
  geom_smooth(method = "lm") +
  # facet_wrap(~ fac_length) +
  facet_wrap(~ cut_number(length, n = 3), ncol = 2) +
  labs(title = "Movie budgets vs. year cut by length")
```

• The budget of movies seem to follow a linear trend with year, hence we will be fitting a linear model to predict budget from year. Furthermore, there is no reason for us to assume that it should follow a curve anyway. Hence, for the sake of simplicity, more than anything else, we will go ahead with *lm()*.

• The dataset does not have any major outliers. Thus, it makes sense to fit a least squares model (no need for a robust fit).

• Conditional plots faceted by length of movies show lines of very different slopes. We will have to account for this interaction in our model. Moreover, coplots cut by year do not showcase anything interesting.

• There does not seem to be any interaction between year and length of movies.

```{r}
# Fit three separate linear models for short, medium and long movies to predict
# budget from year
fac_length = rep(NA, nrow(movies))
fac_length[movies$length < 90] = "short"
fac_length[(movies$length >= 90) & (movies$length < 120)] = "medium"
fac_length[movies$length >= 120] = "long"
movies3 = data.frame(movies, fac_length)

movies.lm = lm(log_budget ~ year * fac_length, data = movies3)
movies.lm.df = augment(movies.lm)
movies.grid = expand.grid(year = seq(1905, 2005, 1), fac_length =
                            c("short", "medium", "long"))
movies.predict = predict(movies.lm, newdata = movies.grid)
movies.grid.df = data.frame(movies.grid, fit = as.vector(movies.predict))
```

# Solution 2

```{r}
# Visualise the model
ggplot(movies.grid.df, aes(x = year, y = fit, group = fac_length, color = fac_length)) +
  geom_line() +
  scale_color_manual(values = cb_palette) +
  ggtitle("Figure 1: Variation in the budget of movies over the years
conditioned on their length") +
  labs(subtitle = "Budgets have been transformed onto log10 scale",
       caption = "Movie length thresholds set at 90 minutes and 120 minutes.",
       color = "Movie length")
```

While the budgets of movies shorter than 90 minutes have *mostly* remained unchanged, budgets of movies longer than 90 minutes have increased sharply over time; more so for movies longer than 120 minutes.

```{r, include = FALSE}
# Plot the residuals
ggplot(movies.lm.df, aes(x = year, y = .resid)) + geom_smooth() +
  facet_wrap(~fac_length) + labs(title = "Residual plot")
```

# Solution 3

```{r}
wireframe(fit ~ year * fac_length, data=movies.grid.df, screen=list(z=-70, x=-90, y=90))
```
The 3D wireframe plot above confirms our observation that the budget for *long* movies has gone up at a much higher rate than that of the *medium* budget movies over the years, whereas the budget for *short* movies has remained pretty much constant.
